---
title: "Lab 4 Comments: Causation in Observational Studies"
subtitle: "Replicating Ferwerda and Miller (2014)"
author: "Your Name Here"
date: "Last Updated `r format(Sys.Date())`"
output:
  html_document:
    theme: journal
    toc: TRUE
    toc_float: TRUE
    toc_depth: 2
    number_sections: TRUE
---



# Overview {-}

In this assignment, we'll walk through the logic and design of [Ferwerda and Miller (2014).](https://pols2580.paultesta.org/files/readings/ferwerda_miller_2014.pdf) 

## Goals

Conceptually, our goal in this lab is to see how scholars might use historical knowledge to make causal claims with observational data.

Specifically, we will see how F&M leverage a claim about how borders are drawn to assess the effects of different types of governing strategies.

Practically, we will continue to develop our statistical skills, introducing some core concepts from base `R`. 

Spefically we will see how we can use:

- `for()` loops to repeat a process like calculating a mean, over multiple variables
- self-defined functions to abstract and generalize repeated tasks
- the `with()` function to avoid having to write out `df$variable` 
- different types of `apply()` functions (namely `sapply()` and `tapply()`) to apply functions to a sets of variables (`sapply()`) and to subgroups within a set of variables (`tapply()`)

These are useful skills that broadly help you write your code more efficiently.  Things like `for()` loops, `functions()` and `apply()` can reduce the amount of copying, pasting and replacing you have to do, which in turn can reduce the amount of errors induced by forgetting to change a variable name, or mistyping a command.

But the first time you see a for loop, or define your own function, it will likely seem a bit abstract, and obtuse.That's ok. The goal is that you have a better, if not perfect, understanding of these concepts which we will use throughout the course. 

# Read Ferwerda and Miller (2014) and answer the following:

- **What's the research question?** 
- **What's the theoretical framework?** 
- **What's the empirical design?** 
- **What's are the main results?** 



# Load the replication data and set up your workspace

Please run the following command to load the data:


```{r}
# Load the data
load(url("https://pols2580.paultesta.org/files/data/04_lab.rda"))
```

And write code to load any packages you think you might need to analyze the data.

```{r libraries}
# Libraries
```


The next commands tell us a little about the dataset. Specifically, there are 1371 observations (rows) and 14 variables (columns)

```{r}
dim(wrk)
names(wrk)


```

Here's a brief codebook to help you along the way


- **south**: 0-1 indicator, 1 if south of demarcation line (Vichy), 0 otherwise (Germany)
- **distancetotrain**: Distance to nearest train station from center of commune
- **communications**: Ordinal variable treated as interval indicating whether a commune had a telephone bureau (3?), a telegraph station (2?), or a post office (1?) or none of these (0?) in 1939
- **hectares_scaled**: number of actively farmed hectares as a proxy for forest cover and urbanization
- **population**: Population from the 1936 census
- **pctright**: Vote shares for the Right from the 1936 national legislative elections 
- **pctleft**: Vote shares for the Left from the 1936 national legislative elections
- **sd_elev250**: The standard deviation of elevation using resampled 250m resolution SRTM elevation data; a proxy for Ruggedness of Terrain
- **near_km**: Distance, in kilometers to the Demarcation line
- **s**: Count of the number of sabotage activities (all attacks against infrastructure (largely railroad and communications)
- **fighting**:: Count of the number of resistance-initiated attacks on German and Vichy personnel
- **area_sqkm**: Area in square kilometers
- **north**: 0-1 indicator, 1 if north of demarcation line (Germany), 0 otherwise (Vichy)
- **nom\_dept**: Name of the department.


-------------------------

# What's the unit of analysis for this study?

**Answer:** 


# Describe the structure of the data

Specifically:

- How many departments are there in the sample? 
- How many communes in each department are north and south of the demarcation line?


```{r}
# Write your code here

```

**Answer:** 



# Describe the study's outcomes of interest and their distributions 


```{r}
# Write your code here




```

**Answer:** 


-------------------------

# What is the "treatment" in this study? 

What variable or variables in the data describe treatment status?



```{r}
# Write your code here

```

**Answer:** 



-------------------------

So one of the implications of this design is that a commune's status relative to the demarcation line (and thus type of government/devolution of power) is "as-if" random. There's no way to really prove this claim, but we can test some of its observable implications. 

Specifically, if the Germans essentially drew the demarcation line at random (at least relative to the local boundaries of specific communes), then the observable characteristics of communes north of the line (under German occupation) and south of the line (under Vichy rule) should be similar. 


# Calculate the means and differences of means for all communes north and south of the demarcation line for the following variables:

- **distancetotrain**: Distance to nearest train station from center of commune
- **communications**: Ordinal variable treated as interval indicating whether a commune had a telephone bureau (3?), a telegraph station (2?), or a post office (1?) or none of these (0?) in 1939
- **hectares_scaled**: number of actively farmed hectares as a proxy for forest cover and urbanization
- **population**: Population from the 1936 census
- **pctright**: Vote shares for the Right from the 1936 national legislative elections 
- **pctleft**: Vote shares for the Left from the 1936 national legislative elections
- **sd_elev250**: The standard deviation of elevation using resampled 250m resolution SRTM elevation data; a proxy for Ruggedness of Terrain
- **near_km**


```{r}
# Write your code here
# You might find something like the following useful

# # List of covariates
# the_vars<-c("distancetotrain","communications","hectares_scaled","population",
#           "pctright","pctleft","sd_elev250","near_km")
# 
# # Holders for means
# north_means<-c()
# south_means<-c()
# 
# #Replace ???? in code below with some useful logical statement
# for(i in the_vars){
#     print(paste("Now summarizing",i))
#     north_means[i]<-mean(wrk[????,i])
#     south_means[i]<-mean(wrk[????,i])
# 
# 
# }


```


**Answer:**




-------------------------

So how should we make sense of these differences? Are they substantively large? Do they provide support for or against a claim that the demarcation lines is "as if" random?

Answering these questions has proved difficult, since different variables are measured on different scales, and it's hard to quantify what big is. 

Last week, we saw that we could assess these differences relative to a claim about the world ("i.e. if there were no difference between theses groups, how likely is it that we would a difference as large as we did").

Today, we'll consider another way of thinking about how big differences are, which is to think about how much the difference might vary. 

We know from our first week on measurement, that things like variance and standard deviations can be used to describe the spread of a random variable's distribution (how much observations tend to vary around their means). 

Means and differences of means are functions of variables, and so we can think about these quantities as having their own distributions and spreads.

So one way to compare the magnitude of the differences you calcuated above, is to standardize each difference by its standard error -- that is, we could divide each difference by the standard deviation of the sampling distribution of that difference (how much we might expect that difference to vary). Don't worry about the terminology right now--we'll cover these concepts in greater detail later in the course. 

For now, we'll use as a rough rule of thumb, that **standardized differences greater than 2 are "big,"** that is, they are unlikely to arise by chance and thus provide evidence against the claim that the demarcation line was drawn as if random.

The code below **defines a function** to calculate these standardized differences and returns the standardized difference (as well as the means, difference of means, and standard error of that difference) for the `distancetotrain` variable.

```{r}

std_diff_fn <- function(x){
  # inputs
  # x: the variable we want to calculate a standardized difference of
  stopifnot(is.character(x))
  
  #1.  Calculate means and difference of means
  mu1 <- with(wrk, mean(wrk[north==1, x],na.rm=T))
  mu0 <- with(wrk, mean(wrk[north==0, x],na.rm=T))
  diff <- mu1 - mu0

  #2. Calculate "standard error" of difference
      # Assuming unequal variance of groups
  sd1 <- with(wrk, sd(wrk[north==1, x],na.rm=T))
  sd0 <- with(wrk, sd(wrk[north==0, x],na.rm=T))
  n1 <- with(wrk, sum(!is.na(wrk[north==1, x])))
  n0 <- with(wrk, sum(!is.na(wrk[north==0, x])))
  se <- sqrt( sd1^2/n1 + sd0^2/n0)

  #3. Standardize difference
  std_diff <- diff/se
  
  #4. Combine Results
  results <- c(North=mu1, South=mu0,Differnece=diff, SE = se, StdDiff = std_diff)
  return(results)
  
}

std_diff_fn("distancetotrain")


```


# Use the `sapply()` function to apply, the `std_diff_fn` to the covariates in the object `the_vars.` 

Evaluate these standardized differences in relation to the authors' claims that the demarcation line can be seen as being as-if randomly assigned.


```{r}
# Write your code here
# hint:
# sapply(some object here, some function here)
```

**Answer:**



-----------------------------

Hmmm... so using our "standardized differences bigger than two" rule of thumb, we became a little bit skeptical about whether the claim of as-if random was justified for the whole sample.

Now we'll explore whether for a subset of the sample, this claims appears more plausible. 

Specifically, we'll look at whether for communes within a narrower window or bandwidth around the demarcation line differences in observed covariates between communes north and south of the line are small, and thus seem consistent with a claim that the line of demarcation is as good as randomly assigned for these communes.

To do so, we'll modify our `std_diff_fn` and create a new function called `new_std_diff_fn`


```{r}
new_std_diff_fn <- function(x, c = max(wrk$near_km,na.rm=T)){
  # inputs
  # x: the variable we want to calculate a standardized difference of
  # c: the cutoff distance from the borderd
  stopifnot(is.character(x))
  stopifnot(is.numeric(c))

  # What's going on here?
  wrk <- wrk[wrk$near_km <= c,]
  
  #1.  Calculate means and difference of means
  mu1 <- with(wrk, mean(wrk[north==1, x],na.rm=T))
  mu0 <- with(wrk, mean(wrk[north==0, x],na.rm=T))
  diff <- mu1 - mu0

  #2. Calculate "standard error" of difference
      # Assuming unequal variance of groups
  sd1 <- with(wrk, sd(wrk[north==1, x],na.rm=T))
  sd0 <- with(wrk, sd(wrk[north==0, x],na.rm=T))
  n1 <- with(wrk, sum(!is.na(wrk[north==1, x])))
  n0 <- with(wrk, sum(!is.na(wrk[north==0, x])))
  se <- sqrt( sd1^2/n1 + sd0^2/n0)

  #3. Standardize difference
  std_diff <- diff/se
  
  #4. Combine Results
  results <- c(North=mu1, South=mu0,Differnece=diff, SE = se, StdDiff = std_diff)
  return(results)
  
}

new_std_diff_fn("distancetotrain", c=10)

```


# Explain how `new_std_diff_fn()` differences from `std_diff_fn()`



**Answer:** 


-------------------------

Great! So this modified function can calculate conditional means for a subset of observations (communes) within `c` kilometers of the demarcation line.


# Does this claim appear more or less plausible within 10 km of the line?

Use  `sapply()` with `new_std_diff_fn()` to calculate the standardized differences for the covariates in `the_vars` for communes within 10 km north or south of the demarcation line (i.e., c=10). 

Evaluate these differences as they relate to F&M's claim of "as-if" randomization. 


```{r}
# Write your code here

```



**Answer:** 


-------------------------

So the identifying assumption seems more plausible for communes closer to the demarcation line. Now let's estimate the "effect" of German rule on resistance activity (sabotage and fighting) for the full sample and for the subset of communes within 10 km of the demarcation line.


# Does devolution reduce resistance?


Calculate the difference in the mean number of sabotage events and fighting events for communes north and south of the demarcation line, for the full sample of communes and the communes within 10 km of the demarcation line. Provide a substantive interpretation of these differences.

You can use logical indexing, or the `new_std_diff_fn` if you'd like. Try to do it both ways to savor the joy of logical indexing and the magic of functions.

```{r}
# Write your code here

```

**Answer:**





-------------------------


# Repeat the calculations above, but this time comparing the differences between communes north and south of the line within the same department. 

Comment on the relative similarity of or differences in the findings from estimating effects for each department compared to pooling all the communes together?

Below is some code to get you started the hard way, you could also try modifying the `new_std_diff_fn`

```{r}
# Here's some code to get you started

# 10k
# Sabotage
# Calculate means by department for communes in the north within 10 km of line
mus_s_north_dep_10k <- with(wrk[wrk$north==1 & wrk$near_km<=10,], tapply(s, nom_dept, mean))

# Calculate means by department for communes in the south within 10 km of line
mus_s_south_dep_10k <- with(wrk[wrk$north==0 & wrk$near_km<=10,], tapply(s, nom_dept, mean))

# Calculate differences within each department
diffs_s_south_dep_10k <- mus_s_north_dep_10k-mus_s_south_dep_10k

# Print differences.
diffs_s_south_dep_10k

###########
# Your turn
###########

# 10k
# Fighting


# Full sample 
# Sabotage


# Full sample 
# Fighting

```



**Answer:** 