---
title: "Lab 10 / Assignment 3 - Final Project Data Setup and Explorations"
author: "Your Group Members Names Here"
date: "Last Updated `r format(Sys.Date())`"
output:
  html_document:
    theme: journal
    toc: TRUE
    toc_float: TRUE
    toc_depth: 2
    number_sections: TRUE
---

```{r setup, include = FALSE, cache = FALSE}
knitr::opts_chunk$set(error = TRUE, 
                      comment = NA, 
                      warnings = FALSE, 
                      errors = FALSE, 
                      message = FALSE, 
                      cache = TRUE,
                      tidy = FALSE)


```



# Overview {-}

In our final lab, you will apply concepts and skills from this course to explore data from the [2020 American National Election Study](https://electionstudies.org/data-center/2020-time-series-study/). Specifically you will

1. Identify an outcome of interest (5-10 minutes)
2. Identify key predictors and covariates (5-10 minutes)
3. Recode your data (20 minutes)
4. Describe your data (20 minutes)
4. Describe your question, expectations, and models (10 minutes)
5. Estimate, present, and interpret your models (20 minutes, Graded Question)

Ideally, each group will pursue a question that interests them. I will also complete these tasks live, so, if you're not feeling confident, you can follow along with me and submit the code I demo in class as your lab for a grade of 85..



# Set up your workspace and load data  {-}

Here's some code to get you started. Feel free to add additional packages as needed

```{r}
# Libraries ----
# Add any additional needed packages below 
library(tidyverse)
library(labelled)
library(texreg)



# Load 2020 NES ----
# Uncomment if nencessary
# library(devtools)
# install_github("jamesmartherus/anesr")

library(anesr)
data("timeseries_2020")

# Rename to df
df <- timeseries_2020


```


# Identify an outcome of interest

1. Open up the [list of variables](https://electionstudies.org/wp-content/uploads/2021/07/anes_timeseries_2020_varlist_20210719.pdf) in the survey

2. Do a command/ctrl+F find to quickly search for keywords of potential outcomes of interest

3. If you have questions about how a variable is measured, how the question was asked, search the variable name (e.g `V201144x`) in the [codebook](https://electionstudies.org/wp-content/uploads/2022/02/anes_timeseries_2020_userguidecodebook_20220210.pdf)

4. Pick one variable that you wish to explore.

Some potential variables of interest:


- `V202143`-`V202187` various post election feeling thermometers
- `V202073` POST: Presidential vote choice
- `V202231x` - `V202248x` attitudes toward immigration
-  `V201246` - `V201262` ideological policy placements of self and parties
- `V201233` Trust in government

Once you've settled on an outcome of interest, please fill in the following:

- **Conceptual Name:**
  - **Raw variable name:**
  - **Recoded variable name:**

For example:

- **Conceptual Name:** Trust in government
  - **Raw variable name:** `V201233`
  - **Recoded variable name:** `dv_trust_in_gov`


# Indentify at least one key predictor of interest, and at least two covariates of interest

Your key predictor of interest should be something you think explains variation in your outcome of interest.

Your additional covariates should be things you think are:

- Correlated with both the outcome and predictor
- Things that might moderate the relationship between your key predictor, and your outcome of interest

Some commonly used covariates:

- **V201507x** PRE: SUMMARY: Respondent age
- **V201511x** PRE: SUMMARY: Respondent 5 Category level of education
- **V202468x** PRE‐POST: SUMMARY: Total (family) income
- **V201600** PRE: What is your (R) sex? [revised]
- **V201549x** PRE: SUMMARY: R self‐identified race/ethnicity
- **V201231x** PRE: SUMMARY: Party ID
- **V201200** PRE: 7pt scale liberal‐conservative self‐placement

In general, if there is a SUMMARY version of the variable, use that.

As you did above, list your key predictor here:

- Conceptual name:
  - Raw variable name:
  - Recoded variable name:


List your covariates here:

- Conceptual Name: 
  - Raw variable name:
  - Recoded variable name:
  
- Conceptual Name: 
  - Raw variable name:
  - Recoded variable name:  


# Recode your data



1. Look at the distributions of your raw variables, perhaps using `table()` or `summary()`

```{r}
# Frequency distribution
# table(df$V201231x)
# # Look at variable and value labels
# var_label(df$V201231x)
# val_labels(df$V201231x)

```


2. Identify values that need to be recode or transformed 
  - For example are there negative values that need to be recoded as NA
  - Do you wish to create a summary measure that is the average of multiple variables, or the difference or two variables?
  
3. Use `%>%` and `mutate()` along with `ifelse()` and `case_when()` to recode your data as needed. 

Below is some pseudo-code to help get you started


```{r ex_recode}
# # Pseudo code
# # Recode DV ----
# df %>%
#   # Recode DV
#   mutate(
#     dv_name = ifelse(V20XXXX < 0, NA, V20XXXX) # Replace Negative Values with NA
#   ) -> df
# 
# # Check recode
# table(df$dv_name, df$V20XXXX, useNA = "ifany")

# # Recode IVs ----
# df %>%
#   # Recode IV
#   mutate(
#     pid_3cat = case_when(
      #   V201231x < 4 ~ "Democrat",
      #   V20XXXX == 2 ~ "Independent"
      #   V201231x > 4 ~ "Republican",
      #   T ~ NA_character_
      #   ),
      # is_dem = ifelse(pid_3cat == "Democrat", 1, 0),
      # is_rep = ifelse(pid_3cat == "Republican", 1, 0),
      # is_ind = ifelse(pid_3cat == "Independent", 1, 0),
# ) -> df 

# # Check recodes
# table(df$pid_3cat, df$V201231x, useNA = "ifany")



```




# Describe your data

Once you've recoded your data:

- Create a table of summary statistics for your outcome, key predictor(s), and covariates

- Present at least one descriptive figure, that illustrates the distribution of your outcome or key predictor, or shows an interesting relationship between variables.

- Interpret your results

## Summary statistics:

Producing a table of summary statistics requires a little foresight. 

Essentially you want to make a data frame where each row is a (numeric) variable, and each column is a statistic (minimum, 25th percentile, median, mean, 75th percentile, max, Number of missing).

To do this, I would:

- create a object called `the_vars` which contains the names (in quotation marks) of the variables you want to summarize.

- Select these variables from your data set. using `df%>%select(all_of(the_vars))`

- Use `%>%pivot_longer()` specifying `cols=select(all_of(the_vars))`, and `names_to` equals `"Variable"` and `values_to = "value"`  to transform this wide dataset into a long dataset

- Then use `%>%group_by(Variable)%>%` and `summarise()` to calculate the statistics for each variable of interest (e.g. `%>%summarise(Mean = mean(value, na.rm=T)))`)

- Save the output to an object called something like `sum_df`

- In a new chunk use `knitr::kable(sum_df) %>% kableExtra::kable_styling()` to format your table. Set `echo=F` in the code chunk head

```{r}
# Summarise data

```


```{r, results="asis"}
# Display results
```


## Descriptive Figures

To create a figure, you'll need to specificy the following

- `data` (e.g. `df %>%`)

- aesthetic mappings, `ggplot(aes(x = predictor, y = outcome))`

- geometries 

  - Univariate: `geom_density()`, `geom_boxplot()` `geom_histogram()`
  
  - Bivariate: `geom_point()` (for a scatterplot), `geom_line()` for a trend.

Once you have a minimal working example, play around with other grammars of graphics:

- `labs()` for custom labels

- `theme_XXX` for custom themes

- `facet_wrap(~group)` to produce the same plot *facetted* by some categorical grouping variable

When you're happy with your figure, save it as object in R (e.g. `fig1 <- df %>% ggplot(aes(predictor, outcome))+geom_point())`. Put that object in its own chunk to display it in your document.

Don't let the perfect be the enemy of the good.


```{r}
# Descriptive figures

```

## Descrptive Interpretation:

Please provide an overview of the data (source, number of observations, unit of analysis).

Describe a typical observation, making reference to the statistics in your summary table.

Offer a substantive interpretation of your descriptive figure(s). What do they tell us about the distribution of a key variable, or the relationship between two variables.


# Summarize your research question, expecations and empirical models


First, summarize your research question and motivation:

> We're interested in how Y varies with X. Specifically, we expect that .... , since ..... To address concerns about omitted variable bias, we will control for A and B, since ....


Next, describe a simple bivariate model between your outcome and key predictor

$$\text{DV name here} = \beta_0 + \beta_1 \text{Key Predictor here}$$
If your expectations are correct, then what should should the sign and signficance be $\beta_1$


Then describe a multiple regression model


$$\text{DV name here} = \beta_0 + \beta_1 \text{Key Predictor here} + \beta_2 \text{Covariate 1 here} ... + \beta_k\text{Covariate k here}$$
Explain how you expect the coefficient for your key predictor to change $(\beta_1)$, if at all

If you're feeling frisky, you might describe additional models (interaction models, models with polynomial terms, etc.), but you need not.


# Estimate the models described above

## Fit the models

```{r models}
# Model 1: Bivariate Model

# Model 2: Multiple Regression
```


## Display the models in a regression table


```{r, table}
# Regression table
```


## Interpet your models

Please write a 1 paragraph summary interpreting your results in terms of both their statistical and substantive significance. Assume your audience is smart, but has never taken POLS 1600. Explain to them what a regression model is, what a standard error, p-value, and/or confidence interval is. How should they interepret the substantive findings of your model. How should they assess the statistical uncertainty around these results?

Perhaps you might reade create a plot of predicted values from a model to help facilitate the substantive interpretation of your results. If so, here's a code chunk for you:

```{r interpetation}
# Additional code chunk to facilitate interpretation of models
```




