---
title: "A3: Initial Analyses"
format:
  html:
    toc: TRUE
    toc_float: TRUE
    toc_depth: 2
    number_sections: TRUE
---

# Overview {.unnumbered}

The broad goal for this assignment is to ensure that you're making progress on your final project.

The final project that you will submit for this class consists of the following:

1. Introduction (5 percent, ~ 4 paragraphs)
2. Theory and Expectations (10 percent, ~4+ paragraphs)
3. Data (20 percent ~ 4+ paragraphs)
4. Design (25 percent ~ 5+ paragraphs)
5. Results (25 percent ~ 5+ paragraphs)
6. Conclusion (5 percent ~ 3+ paragraphs)
7. Appendix (10 percent ~ Variable codebook and all the R code for your project)

The specific expectations for each section will be discussed in further detail in your next assignment. 

Next week's assignment is [linked here](https://josephessig-source.github.io/pols1600_site/assignments/a4.html), which may help with your outlining. 

Our focus this week will be on the logistical aspects of your final project:

- <a href="https://josephessig-source.github.io/pols1600_site/files/pols1600_paper_template.qmd" download="" >  Downloading the template </a> for your final paper for your final paper
- Making sure you're set up for collaboration with GoogleDrive
- Finish wrangling your data in the code of your final paper
  - Loading the data into R
  - Inspect the data
  - Completing recoding and if necessary, merging of the data
- Begin describing the data
- Begin outlining the other sections of your document (Intro, Theory, Data & Design, etc.)

Next week, we will focus on estimating, interpreting and presenting your results.

# Initial Analyses

## Data wrangling and inspection

Load your data into R and start getting a HLO of the data. Subset the data to just the variables you will need. 

```{r}

```

## Recoding

In this chunk, work on creating any new variables or making any transformations of variables you will need, like creating a factor, assigning a numerical value based on labels, etc. 

```{r}

```


## Describing the data

Provide descriptive statistics for some of your variables. Create at least 1 figure to visualize your data (it does not have to be pretty, or be the final version you will include in your paper.)

```{r}

```


## Outlining the paper 

Please provide a few sentences or bullet points giving a broad overview of each section for your paper. 

### Introduction

- What is your research question? 

### Theory

- What is the phenomena you are studying? (Research Question/Outcome of Interest)

- Why should we care about this phenomena? (Context/Motivation/Importance)

- What are the factors scholars think explain this phenomena? (Literature Review/Key predictors/Alternative explanations). Identify at least 1 relevant paper.


### Data

- What is your data, and where does it come from? Is the data approriate for your RQ? 

- What is the unit of analysis? 


### Design

- How will you test your theoretical expectations?

- Describe, in words, at least 1 regression model you will run. 

- Results

- What do you expect your results to look like? You don't need to go into detail on the specifics of your analysis, just think about the direction of your potential results.





