% !TEX root = ./POLS_1600_syllabus_Spring_2025.tex


\part*{Schedule}


\end{introstuff}

\SetDate[23/01/2025]

The general outline of topics for the course is as follows. The data and topics for some labs may change.  

In general, you will 

\begin{itemize}
\item Do the week's assigned reading before class on Tuesday
\item Submit your labs in class on Thursday
\item Upload your weekly QSS Tutorial problem by midnight, Friday.
\end{itemize}

\section{Week 0 --- \today --- Introduction and Course Overview}

 
\subsection{Thursday} \textbf{\today---Lecture:} What am I getting myself into? Course overview
\AdvanceDate[5]

\subsection{Read:}

\citet{Imai2022-pm} Chapter 1 (if you have the textbook)

\subsection{Do:}

Download and install R and R Studio following the steps outlined in \href{https://pols1600.paultesta.org/labs/00-software_setup.html}{Getting started with R}

\vspace{2em}

\section{Week 1. Describing Data in R}

\subsection{Read:}

\citet{Imai2022-pm} Finish Chapter 1 
\citet{Imai2022-pm} Chapter 3 pp. 88-105 


\subsection{Tuesday} \textbf{\today---Lecture:} How do we describe the world around us?
\AdvanceDate[2]

\subsection{Thursday} \textbf{\today---Lab:} Levels of measurement; Measures of central tendency and dispersion; Loading and manipulating data in R
\AdvanceDate[1]
\subsection{Friday} \textbf{\today---Tutorials:} 
QSS Tutorial 0: Introduction to R; QSS Tutorial 1: Measurement 1
\AdvanceDate[4]



\vspace{2em}


\section{Week 2. Visualizing Data in R}

\subsection{Read:}

\citet{Imai2022-pm} Chapter 3 pp. 105-141
R 4 Data Science Chapter 3 Data Visualization: \url{https://r4ds.had.co.nz/data-visualisation.html}


\subsection{Tuesday} \textbf{\today---Lecture:} How can we visualize data to help us understand the world around us?
\AdvanceDate[2]

\subsection{Thursday} \textbf{\today---Lab:} Visualizing distributions and relationships with ggplot
\AdvanceDate[1]
\subsection{Friday} \textbf{\today---Tutorial:} QSS Tutorial 2: Measurement 2
\AdvanceDate[4]



\vspace{2em}



\section{Week 3. Causation I - Experimental Studies}

\subsection{Read:}

\citet{Imai2022-pm} Chapter 2 pp. 38-65

\citet{Broockman2016-pi} skim for understanding of basic question, data and design



\subsection{Tuesday} \textbf{\today---Lecture:}  Making Causal Claims in Experimental Studies: How do we know if X causes Y?
\AdvanceDate[2]

\subsection{Thursday} \textbf{\today---Lab:} Estimating causal effects with experimental data
\AdvanceDate[1]
\subsection{Friday} \textbf{\today---Tutorial:} QSS Tutorial 3: Causation 1
\AdvanceDate[4]


\vspace{2em}

\section{Week 4. Causation II - Observational Studies}

\subsection{Read:}

\citet{Imai2022-pm} Chapter 2 pp. 65-88


\subsection{Tuesday} \textbf{\today---NO CLASS:} 


\AdvanceDate[2]

\subsection{Thursday} \textbf{\today---Lecture:}  Making Causal Claims in Observational Studies: How do we know if X causes Y without randomly assigning X? 

\AdvanceDate[1]
\subsection{Friday} \textbf{\today---Tutorial:} QSS Tutorial 4: Causation 2
\AdvanceDate[2]

\vspace{2em}

\subsection{Assignment 1} \textbf{DUE: 5 pm on Sunday, \today. Submit proposed research topics on Canvas}

\AdvanceDate[2]


\vspace{2em}

\section{Week 5. Prediction I}

\subsection{Read:}

\citet{Imai2022-pm} Chapter 4 pp. 144-188


\subsection{Tuesday} \textbf{\today---Lecture:} How do we make predictions? Simple linear regression. OLS as a linear approximation to the conditional mean function. 
\AdvanceDate[2]

\subsection{Thursday} \textbf{\today---Lab} Using simple linear regression to explore the phenomena of Red Covid 

\AdvanceDate[1]
\subsection{Friday} \textbf{\today---Tutorial:} QSS Tutorial 5: Prediction 1
\AdvanceDate[4]




\vspace{2em}

\section{Week 6. Prediction II}


\subsection{Read:}

\citet{Imai2022-pm} Chapter 4 pp. 188-214


\subsection{Tuesday} \textbf{\today---Lecture:} How do we make predictions adjusting for potentially confounding factors? Multiple regression. What it means to ``control for education'' or hold ``age constant''. 

\AdvanceDate[2]

\subsection{Thursday} \textbf{\today---Lab:} Exploring alternative explanations for Red Covid

\AdvanceDate[1]
\subsection{Friday} \textbf{\today---Tutorial:} QSS Tutorial 6: Prediction 2
\AdvanceDate[2]

\vspace{2em}

\subsection{Assignment 2} \textbf{DUE: 5 pm on Sunday, \today. Submit discussion of potential data sources to Canvas}

\AdvanceDate[2]

\vspace{2em}


\section{Week 7. Prediction III}

\subsection{Read:}

Review \citet{Imai2022-pm} Chapter 4

Read \citet{Grumbach2022-zj}

\subsection{Tuesday} \textbf{\today---Lecture} How do we evaluate our models predictions?

\AdvanceDate[2]

\subsection{Thursday} \textbf{\today---Lab:} Understanding the data and design of \citet{Grumbach2022-zj}


\AdvanceDate[5]

\vspace{2em}

\section{Week 8. Probability I}

\subsection{Read:}

\citet{Imai2022-pm} Chapter 6 pp. 279-342

\subsection{Tuesday} \textbf{\today---Lecture:} What is probability and how do we use it? Axioms of probability; Discrete and continuous probability distributions; Expectations, variance, and moments 
\AdvanceDate[2]

\subsection{Thursday} \textbf{\today---Lab}: Replicating the results of \citet{Grumbach2022-zj}

\AdvanceDate[1]
\subsection{Friday} \textbf{\today---Tutorial:} QSS Tutorial 7: Probability 1
\AdvanceDate[2]


\vspace{2em}
\subsection{Assignment 3} \textbf{DUE: 5 pm on Sunday, \today. Submit Data Explorations on Canvas}

\AdvanceDate[2]
\AdvanceDate[7]


\vspace{2em}

\part*{Spring Break}

\vspace{2em}


\section{Week 9. Probability II}

\subsection{Read:}

\citet{Imai2022-pm} Chapter 6 pp. 342-354


\subsection{Tuesday} \textbf{\today---Lecture:} What is probability and how do we use it? Conditional probability; Bayes Rule; The Law of Large Numbers and The Central Limit Theorem;
\AdvanceDate[2]

\subsection{Thursday} \textbf{\today---Lab:} Exploring data from the 2024 \href{http://electionstudies.org/}{National Election Study}

\AdvanceDate[1]
\subsection{Friday} \textbf{\today---Tutorial:} QSS Tutorial 8: Probability 2
\AdvanceDate[4]

\section{Week 10. Uncertainty I}

\subsection{Read:} 
\citet{Imai2022-pm} Chapter 7 pp. 357-417

\subsection{Tuesday} \textbf{\today---Lecture:} How do we quantify uncertainty? Asymptotic and simulation based approaches to sampling distributions, standard errors, and confidence intervals. 

\AdvanceDate[2]

\subsection{Thursday} \textbf{\today---Lab} Exploring data from the 2024 \href{http://electionstudies.org/}{National Election Study}

\AdvanceDate[1]
\subsection{Friday} \textbf{\today---Tutorial:} QSS Tutorial 9: Uncertainty 1
\AdvanceDate[4]



\vspace{2em}


\section{Week 11. Uncertainty II}

\subsection{Read:}

\citet{Imai2022-pm} Chapter 7 pp. 418-445

\subsection{Tuesday} \textbf{\today---Lecture:} How do we quantify uncertainty? Asymptotic and permutation based approaches to hypothesis test and p-values 
\AdvanceDate[2]

\subsection{Thursday} \textbf{\today---Lab} Workshopping your final papers
\AdvanceDate[1]
\subsection{Friday} \textbf{\today---Tutorial:} QSS Tutorial 10: Uncertainty 2
\AdvanceDate[2]

\vspace{2em}
\subsection{Assignment 4} \textbf{DUE: 5 pm on Sunday, \today. Submit draft of final project to Canvas}

\AdvanceDate[2]


\vspace{2em}


\section{Week 12. Workshops }


\subsection{Tuesday} \textbf{\today---Workshop:} Course Review and Final Project Workshop
\AdvanceDate[2]

\subsection{Thursday } \textbf{\today---Workshop:} Course Review and Final Project Workshop

\AdvanceDate[3]

\vspace{2em}



\subsection{Assignment 5} \textbf{DUE: 5 pm on Sunday, \today. Upload Project Presentations To Canvas}
\AdvanceDate[2]
\vspace{2em}

\section{Week 13. Workshops }


\subsection{Tuesday} \textbf{\today---Presentations:} Class Presentations
\AdvanceDate[2]

\subsection{Thursday } \textbf{\today---Presentations:} Class Presentations

\AdvanceDate[3]

\vspace{2em}

\subsection{Final Project} \textbf{DUE: 5 pm on Sunday, \today. Upload Final Project To Canvas}

\vspace{2em}



% \section{Week 13. Explorations and Final Paper Workshops}

% \subsection{Read:}

% \citet{Imai2022-pm} Chapter 5

% \subsection{Monday} \textbf{\today---Lecture:}  Data visualization. 
% \AdvanceDate[2]

% \subsection{Wednesday} \textbf{\today---Workshop} Troubleshooting and fine-tuning your drafts
% \AdvanceDate[2]

% \subsection{Friday} \textbf{\today---Workshop} Troubleshooting and fine-tuning your drafts 


% \AdvanceDate[3]

% \vspace{2em}

% \section{Week 14. Explorations and Final Paper Workshops}

% \subsection{Read:}

% \citet{Imai2022-pm} Chapter 5

% \subsection{Monday} \textbf{\today---Lecture:} General feedback on drafts.

% \AdvanceDate[2]


% \subsection{Wednesday} \textbf{\today---Peer review}  You'll be asked to comment on two of your peers' drafts.

% \AdvanceDate[2]

% \subsection{Friday} \textbf{\today---Last Class:} Some parting thoughts and last minute questions. Maybe pizza. Or bagels. Or perhaps pizza bagels.


% \AdvanceDate[7]

% \vspace{1in}

% \part*{Final Papers due by 11:59 PM, Friday, \today \hspace{.1em}  on Canvas}

% \clearpage



