---
title: 'Lab 03 - Replicating Broockman and Kalla (2016)'
author: "Your Group Members Names Here"
date: "Last Updated `r format(Sys.Date())`"
format:
  html:
    toc: true
    toc-location: right
    toc-float: true
    toc-depth: 2
    number-sections: true
---

```{r}
#| label: setup
#| include: false
#| cache: false
knitr::opts_chunk$set(error = TRUE, 
                      comment = NA, 
                      warnings = FALSE, 
                      errors = FALSE, 
                      messages = FALSE, 
                      cache = TRUE,
                      tidy = FALSE)

```

# Overview {.unnumbered}

Today we will explore **the logic and design** of Broockman and Kalla's 2016 study, ["Durably reducing transphobia: A field experiment on door-to-door canvassing"](https://www.science.org/doi/10.1126/science.aad9713){target="_blank"}, from the recruitment of subjects for the study to the delivery of their interventions. Then we will explore **whether the intervention had any effect on respondents' feelings toward transgender individuals.**

To accomplish this we will:

1.  Summarize the study (5 Minutes)
2.  Set up our work space (2-3 Minutes)
3.  Load a portion of the replication data (1-2 Minutes)
4.  Get a high level overview of the data (5 minutes)
5.  Examine the difference in covariates between those assigned to each treatment condition in the study. (10 minutes)
6.  Estimate the average treatment effect of the intervention (10 minutes)
7.  Plot the results and comment on the study (10 minutes)
One of these 7 tasks  will be randomly selected as the graded question for the lab.

You will work in your assigned groups. Only one member of each group needs to submit the html file of lab.

This lab **must** contain the names of the group members in attendance.

# Goals {.unnumbered}

Conceptually, this lab will give you lots of practice calculating **means** and **conditional means**. We will use these means to

-   **Assess the empirical implications of the identifying assumptions of this design** If treatment was randomly assigned, then participants assigned to treatment condition should on average look similar to participants in the control condition (Question 5)
-   **Estimate the average treatment effect over multiple time periods** (question 6). You'll recall from lecture that the ATE is simply a difference of conditional means:

$$
\begin{align*}
E \left[ \frac{\sum_1^m Y_i}{m}-\frac{\sum_{m+1}^N Y_i}{N-m}\right]&=\overbrace{E \left[ \frac{\sum_1^m Y_i}{m}\right]}^{\substack{\text{Average outcome}\\
\text{among treated}\\ \text{units}}}
-\overbrace{E \left[\frac{\sum_{m+1}^N Y_i}{N-m}\right]}^{\substack{\text{Average outcome}\\
\text{among control}\\ \text{units}}}\\
&= E [Y_i(1)|D_i=1] -E[Y_i(0)|D_i=0]
\end{align*}
$$

We'll see that these quantities can be calculated quickly for multiple variables using the following commands:

-   We will use the `group_by` and `summarise` commands to quickly calculate average values for different groups, and the `mutate` function to calculate differences between these averages
-   We will introduce the `across()` and `starts_with()` functions to calculate summaries like a `mean` across multiple variables that start the the same pre-fixes.
-   We will introduce the `pivot_longer()`, `pivot_wider()` and `left_join()` commands to gather, spread, and merge data so that we can calculate the difference of means simply by subtracting means for one group in one column from means for another group in a different column
-   We will use `ggplot()` and related functions to graphically display our results.

# Please render this .qmd file {.unnumbered}

As with every lab, you should:

-   Download the file
-   Save it in your course folder
-   **Update the `author:` section of the YAML header to include the names of your group members in attendance.**
-   Render the document (Check `[X] Render on Save` if desired)
-   Open the html file in your browser (Easier to read)
-   Render the document again after completing a section or chunk (Error checking)
-   Upload the final lab to [Canvas](https://canvas.brown.edu/courses/1102109/assignments){target="_blank"}.

# Summarize the study

Before we get started working with the data, please provide some brief answers (1-3 sentences) to the following questions

-   **What's the research question?** Broadly the Broockman and Kalla (2016) is interested in questions about persuasion. How do we change peoples minds about controversial political topics. Specifically, they are interested in whether a brief conversation in which participants are given information on transgender rights and encouraged to engage in **analogic perspective** taking can lead to lasting changes in attitudes on these issues.
-   **What's the empirical design?** The authors conducted a randomized placebo controlled design. Participants who were registered voters in Miami-Dade County were recruited to take a baseline survey. Among those who completed the baseline survey, subjects were randomly assigned either the treatment or the placebo control condition. In the treatment condition, canvassers from a local organization would knock on subjects doors and deliver a script that provided information about transgender issues and encouraged respondents to try to look at these issues from the perspective of a transgender person. In the placebo control condition, subjects received a message about recycling. Including a "placebo" control condition allows the authors to make comparisons between more similar groups (i.e. people who would actually answer their door when a canvasser knocked). The study also contained a number of post-treatment interventions. 
-   **What's are the main results?** Broockman and Kalla find that these brief 10-minute conversations increased tolerance toward transgendered indviduals by roughly a third of standard deviation and these effects persisted long after the intervention. These changes in general attitudes appear to in turn increase support for non-discrimination laws.

# Set up your workspace

Here is some minimal code so that your lab will compile when you first knit it.

`# Set up your workspace

Here is the code to set up your workspace. 

```{r}
#| label: packages
#| echo: true

library(tidyverse)
library(lubridate)
library(forcats)
library(haven)
```


------------------------------------------------------------------------

# Load the replication data

Next we'll load the data for today, which is a subset of Broockman and Kalla's full data set[^1]

[^1]: You can find the full set of replication files [here](https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/WKR39N){target="_blank"}

1.  **Please run the code chunk below to load the data** That's it. That's the task for this question

```{r}
#| label: data

load(url("https://josephessig-source.github.io/pols1600_site/files/data/03_lab.rda"))
```

The will load a data frame called `df` containing the following variables:

-   `completed_baseline` whether someone completed the baseline survey ("Survey") or not ("No Survey")
-   `treatment_assigned` what intervention someone who completed the baseline survey was assigned two (treatment= "Trans-Equality", placebo = "Recycling")
-   `answered_door` whether someone answered the door ("Yes") or not ("No") when a canvasser came to their door
-   `treatment_group` the treatment assignments of those who answered the door and the received the treatment or placebo (treatment= "Trans-Equality", placebo = "Recycling")
-   `vf_age` the age of the person in years
-   `vf_female` the respondent's sex (female = 1, male = 0)
-   `vf_democrat` whether the person was a registered Democract (Democrat=1, 0 otherwise)
-   `vf_white` whether the person was white (White=1, 0 otherwise)
-   `vf_vg_12` whether the person voted in the 2012 general election (voted = 1, 0 otherwise)
-   `therm_trans_tX` a feeling thermometer,[^2] where respondents are asked to rate how warmly they feel toward transgender people on a scale of 0 (Very Cold) to 100 (Very Warm), with 50 corresponding to neither warm nor cold. The `_tX` corresponds to the time the measure was taken:
    -   `_t0` pre-treatment measure from baseline survey
    -   `_t1` post-treatment measure 3-days after intervention
    -   `_t2` post-treatment measure 3-weeks after intervention
    -   `_t3` post-treatment measure 6-weeks after intervention
    -   `_t4` post-treatment measure 3-months after intervention

[^2]: The actual study contains a number of measures about transgender attitudes and policies which are scaled together to produce a single measure of subjects latent tolerance. For simplicity, we'll focus on this single survey item.

# Provide a high level overview of the data

In the code chunk below, please write code to provide a *high level overview* of the data that allows you to answer the following questions.

Please interpret the results of your code by writing brief answers in Markdown after each question.

1.  **How many observations are there in `df`?** There are `r scales::comma(dim(df)[1])` observations in `df`
2.  **How many variables are there in the `df`?** There are `r dim(df)[2]` variables in `df`
3.  **What's the unit of analysis?** Each row corresponds to a registered voter from Miami Dade County who was invited to take the baseline survey.
4.  **How many registered voters completed the baseline survey?** `r scales::comma(table(df$completed_baseline)[2])` registered voters took the baseline survey or `r scales::percent(prop.table(table(df$completed_baseline))[2])` of the registered voters who were invited to take the survey
5.  **How many treatment conditions are there?** There are two treatment conditions, an intervention in which canvassers

> informed voters that they might face a decision about the issue (whether to vote to repeal the law protecting transgender people); can- vassers asked voters to explain their views; and canvassers showed a video that presented arguments on both sides. Canvassers also defined the term “transgender” at this point and, if they were trans- gender themselves, noted this. The canvassers next attempted to encourage “analogic perspective- taking” (16). Canvassers first asked each voter to talk about a time when they themselves were judged negatively for being different. The can- vassers then encouraged voters to see how their own experience offered a window into transgen- der people’s experiences, hoping to facilitate voters’ ability to take transgender people’s perspectives. The intervention ended with another attempt to encourage active processing by asking voters to describe if and how the exercise changed their mind. (Broockman and Kalla 2016, p. 221)

And a placebo condition in which canvassers instead delivered a message about recycling.

6.  **How many participants were assigned to each condition?** `r scales::comma(table(df$treatment_assigned)[2])` participants who completed the baseline survey were randomly assigned to receive the intervention about Trans Equality and `r scales::comma(table(df$treatment_assigned)[1])` received the placebo message about Recycling.
7.  **How many participants answered the door when a canvasser came?** `r scales::comma(table(df$answered_door)[2])` registered voters actually answered the door when canvassers knocked and `r scales::comma(table(df$answered_door)[1])` did not.
8.  **How many participants actually received each treatment condition?** `r scales::comma(table(df$treatment_group)[2])` registered voters received the Trans-Equality treatment and `r scales::comma(table(df$treatment_group)[1])` received the placebo message about recycling.

```{r}
#| label: hlo

# 1. How many observations are there in `df`?
dim(df)[1]

# 2. How many variables are there in the `df`?
dim(df)[2]


# 3. What's the unit of analysis?
head(df)

# 4. How many registered voters completed the baseline survey?
table(df$completed_baseline)
# 5. How many treatment conditions are there? 
length(unique(na.omit(df$treatment_assigned)))
# OR
table(df$treatment_assigned)


# 6. How many participants were assigned to each condition?
table(df$treatment_assigned)

# 7. How many participants answered the door when a canvasser came?
table(df$answered_door)

# 8. How many participants actually received the treatment?
table(df$treatment_group)



```


# Examine the difference in covariates between those assigned to each treatment condition in the study.

One of the identifying assumptions of an experimental design like Broockman and Kalla's is that treatment is independent of potential outcomes and covariates.

One of the empirical implications of this assumption is that, that the differences in observable covariates between treatment and control should be small.

To the extent the groups are comparable on observable covariates, we are more confident that the random assignment of treatment conditions has created a credible counterfactual comparison.

#| label: baltreat

```{r}
df %>%
  filter(!is.na(treatment_assigned)) %>% # Keep only those assigned a treatment
  group_by(treatment_assigned) %>%
  summarise(across(starts_with("vf_"), \(x) mean(x, na.rm=TRUE))) %>%
  pivot_longer(
    cols = starts_with("vf_"),
    values_to = "Means"
  ) %>%
  pivot_wider(
    names_from = treatment_assigned,
    values_from = Means
  ) %>%
  mutate(
    Diff = `Trans-Equality` - Recycling
  )
```

1.  **Please describe the covariate differences between each treatment condition. Are these differences large or small? **

Subjects assigned to the treatment condition tended to be a little older (by about 1.3 years), were a little less likely to be women (by 1 percentage point) and have voted in the 2012 election (by about 3.7 percentage points) and a little more likely to be Democrats (by 2.4 percentage points) and white (0.7 percentage points).

These differences are smaller than the differences between those who did and did not complete the baseline survey.

A more formal test of these differences show none rise to conventional levels of significance (although some are close).

In short, randomization appears to have succeeded in creating groups that are balanced on observable covariates

```{r}
#| label: baltreat1

df %>%
  # Get the select the columns whose names start with vf_
  select(starts_with("vf_"))%>%
  # Extract the column names
  names()%>%
  # Create the formulas to calculate the diff in means for
  # Each variable
  purrr::map(~ formula(paste0(.," ~ treatment_assigned"))) %>%
  # Caclulate the difference in means for each variable
  purrr::map(~difference_in_means(., data = df)) %>%
  # Put the output into a tidy data frame.
  map_df(tidy)
```

Among those who actually answered the door when canvassers came, we see that that those who received the Trans Equality treatment tended to be less likely to have voted in the 2012 election and this difference seems unlikely to have occurred just by chance (p \< 0.05).

```{r}
#| label: baltreat2

df %>%
  # Get the select the columns whose names start with vf_
  select(starts_with("vf_"))%>%
  # Extract the column names
  names()%>%
  # Create the formulas to calculate the diff in means for
  # Each variable
  purrr::map(~ formula(paste0(.," ~ treatment_group"))) %>%
  # Caclulate the difference in means for each variable
  purrr::map(~difference_in_means(., data = df)) %>%
  # Put the output into a tidy data frame.
  map_df(tidy)
```

If there there large differences across multiple covariates, we might worry that something systematic was going on with the canvassers that made participants more or less likely to answer the door.

Even if we're confident these differences aren't something systematic, we can, as Broockman and Kalla do, adjust our estimates to control for these small differences to ensure that we're isolating the effect of the treatment and not conflating this with other chance variations the voting histories or demographics of the subjects in the treatment and the control.

Statistically significant differences do happen by chance: 1 time out of 20 we'd expect a p-value below 0.05 even when there was no real difference.

The probability of a false positive (i.e. of detecting a difference that isn't there) increases as we make more comparisons (test multiple differences)

We can [adjust our p values for these multiple](https://benwhalley.github.io/just-enough-r/multiple-comparisons.html){target="_blank"} to control for this increased probablity of making false discoveries. When we do, the differences in voting rates are no longer statistically significant.

```{r}
#| label: baltreat3

df %>%
  # Get the select the columns whose names start with vf_
  select(starts_with("vf_"))%>%
  # Extract the column names
  names()%>%
  # Create the formulas to calculate the diff in means for
  # Each variable
  purrr::map(~ formula(paste0(.," ~ treatment_group"))) %>%
  # Caclulate the difference in means for each variable
  purrr::map(~difference_in_means(., data = df)) %>%
  # Put the output into a tidy data frame.
  map_df(tidy)%>%
  mutate(
    p.value = p.adjust(p.value, method = "BH")
  )
```

# Calculate the Average Treatment Effect

Now let's take the same approach to calculate the Average Treatment Effect (ATE) for each time period.

Using the code from the previous sections as a template,

-   Add a `filter(!is.na(treatment_group))%>%` after `df%>%` to filter out subjects who didn't receive the treatment.[^3]
-   Change `summarise(across(starts_with("vf_"), \(x) mean(x, na.rm=T))%>%` code to select variables that start with `therm_trans`
-   Change the `cols` argument in `pivot_longer()` to select the columns that start with `therm_trans`
-   Everywhere you see `treatment_assigned` in the code replace it with `treatement_group`
-   In the `mutate()` function, change:
    -   `Diff_Treatment` to `ATE`
    -   Add the following line `Days = c(0,3,21,42,90)` to create a column called Days (\*make sure to put a comma after `ATE =`Trans-Equality`- Recycling,` ) Be sure to wrap `Trans-Equality` in backticks \` before and after the phrase
-   Change `-> cov_treatment` to `ate_df`

[^3]: Recall that only some people who completed the baseline and were assigned to receive the treatment actually answered the door when canvassers came knocking.

```{r}
#| label: ate

df %>%
  # Calculate difference between Survey and Non-Survey Takers
  group_by(treatment_assigned) %>%
  filter(!is.na(treatment_assigned)) %>%
  # Calculate means for voter file variables
  summarise(across(starts_with("therm_trans"), \(x) mean(x, na.rm=T)))%>%
  pivot_longer(
    cols = starts_with("therm_trans"),
    names_to = "Covariate",
    values_to = "Means"
  ) %>%
  pivot_wider(
    names_from = treatment_assigned,
    values_from = Means
  )%>%
  # Calculate difference
  mutate(
    ATE = `Trans-Equality` - Recycling,
    Days = c(0,3,21,42,90)
    )-> ate_df

ate_df
```

1.  **Did the treatment seem to effect respondents feelings toward transgender people?** The treatment appears to increase respondents favorable feelings toward transgender individuals by about 4 to 6 points
2.  **Does this effect appear to last?** The effects appear relatively stable and persistent, as 90 days after the intervention, respondents who got the Trans-Equality treatment still rated transgender individuals more about 5.3 points higher than subjects who got the Recycling Placebo.

# Plot the results and comment on the study

Using `ate_df` with commands from the `ggplot` package please produce a figure that displays these differences

-   You'll need to provide
    -   `data`
    -   `aes`
    -   `geom`

Here's a simple plot visualzing the ATE's

```{r}
#| label: plotate

ate_df %>%
  ggplot(aes(Days, ATE))+
  geom_point()+
  geom_hline(yintercept = 0,linetype=2)+
  theme_bw()+
  labs(x="Days Since Treatment",
       y = "ATE on Feeling Thermometers\n toward Transgender People")
```

If we wanted to produce something more simlar to Figure 1, we'd need to calculate **confidence intervals** around our estimate. Again, getting ahead of ourselves, a confidence interval is a coverage interval with the property that some $\alpha$ percent of intervals constructed in this manner will contain the true value of quantitantity of interest. When we construct a 95% confidence interval, we are saying that 95 percent of the time this interval contains the "truth" (5 percent of the time the truth lies outside this interval). So we think of confidence intervals as conveying a *range of plausible* values for what the true difference between people who got the treatment and people who got the placebo might be. When 0 lies within this range, we conclude that the difference is "non-significant" since the estimate could have been negative or could have been positive (both are equally likely in a [Frequentist](https://en.wikipedia.org/wiki/Frequentist_inference) view of confidence intervals).

Below we see that the ATE appears non significant 21 days and 90 days after the intervention. The confidence intervals reported by Broockman and Kalla in figure 1 are generally much farther from 0. They're using multiple measures to capture a more general measure of transgender tolerance. In general, using multiple measures of a complex concept [reduces measurement error](https://www.cambridge.org/core/journals/american-political-science-review/article/abs/strength-of-issues-using-multiple-measures-to-gauge-preference-stability-ideological-constraint-and-issue-voting/16F6AF97F7B71AA0112EC9ADF78B553A){target="_blank"}, resulting in more precise estimates.

```{r}
#| label: atecis

df %>%
  # Get the select the columns whose names start with vf_
  select(starts_with("therm_trans"))%>%
  # Extract the column names
  names()%>%
  # Create the formulas to calculate the diff in means for
  # Each variable
  purrr::map(~ formula(paste0(.," ~ treatment_group"))) %>%
  # Caclulate the difference in means for each variable
  purrr::map(~difference_in_means(., data = df)) %>%
  # Put the output into a tidy data frame.
  map_df(tidy)%>%
  mutate(
    ATE = estimate,
    Days = c(0,3,21,42,90)
  ) -> ate_df

ate_df %>%
  ggplot(aes(Days, ATE))+
  geom_point()+
  geom_linerange(aes(ymin=conf.low, ymax=conf.high))+
  geom_hline(yintercept = 0,linetype=2)+
  theme_bw()+
  labs(x="Days Since Treatment",
       y = "ATE on Feeling Thermometers\n toward Transgender People")
```

1.  **Finally, use this space to share some thoughts on the study**

-   **Do you find it convincing?** I do.
-   **What makes the design credible?** Random assignment paired with a placebo control.
-   **Are there things you would do differently?** I do wonder a bit about attrition across multiple waves (some people who completed the baseline survey don't respond to future surveys), and what the proper way scale their multiple measures into a single outcome measure is. There aren't huge differences on our observable covariates between those who participated in the baseline and those who participated in the final wave, so maybe it doesn't make a huge difference, but in theory, you could try and bound the results, imputing extreme values to those who drop out, to get a sense of how much attrition might change the conclusions.

```{r}
#| label: attrition

sum(is.na(df$therm_trans_t0[!is.na(df$treatment_group)]))
sum(is.na(df$therm_trans_t1[!is.na(df$treatment_group)]))
sum(is.na(df$therm_trans_t2[!is.na(df$treatment_group)]))
sum(is.na(df$therm_trans_t3[!is.na(df$treatment_group)]))
sum(is.na(df$therm_trans_t4[!is.na(df$treatment_group)]))

df %>%
  filter(!is.na(treatment_group))%>%
  filter(!is.na(therm_trans_t0))%>%
  group_by(treatment_group)%>%
   # Calculate means for voter file variables
  summarise(across(starts_with("vf_"), mean))%>%
  # Gather columns containing means, exclude column for groups
  pivot_longer(
    cols = starts_with("vf_"),
    names_to = "Covariate",
    values_to = "Means"
  ) %>%
  pivot_wider(
    names_from = treatment_group,
    values_from = Means
  )%>%
  # Calculate difference
  mutate(
    Diff_t0 = `Trans-Equality` - Recycling
  ) -> cov_bal_t0

df %>%
  filter(!is.na(treatment_group))%>%
  filter(!is.na(therm_trans_t4))%>%
  group_by(treatment_group)%>%
   # Calculate means for voter file variables
  summarise(across(starts_with("vf_"), mean))%>%
  # Gather columns containing means, exclude column for groups
  pivot_longer(
    cols = starts_with("vf_"),
    names_to = "Covariate",
    values_to = "Means"
  ) %>%
  pivot_wider(
    names_from = treatment_group,
    values_from = Means
  )%>%
  # Calculate difference
  mutate(
    Diff_t4 = `Trans-Equality` - Recycling
  ) -> cov_bal_t4

cov_bal_t0
cov_bal_t4
```

-   **Do you think the results would "travel" to other topics and issues?** I think it would be interesting to apply this approach to questions of racial justice or partisan polarization.

## Final Trivia

If you're interested, the intellectual history of this study is quite the tale:

-   [The Case of the Amazing Gay-Marriage Data: How a Graduate Student Reluctantly Uncovered a Huge Scientific Fraud](https://www.thecut.com/2015/05/how-a-grad-student-uncovered-a-huge-fraud.html)
-   [How a Gay-Marriage Study Went Wrong](https://www.newyorker.com/science/maria-konnikova/how-a-gay-marriage-study-went-wrong)
-   [How Two Grad Students Uncovered An Apparent Fraud — And A Way To Change Opinions On Transgender Rights](https://fivethirtyeight.com/features/how-two-grad-students-uncovered-michael-lacour-fraud-and-a-way-to-change-opinions-on-transgender-rights/)
-   [How to talk someone out of bigotry](https://www.vox.com/2020/1/29/21065620/broockman-kalla-deep-canvassing)
