---
title: 'Lab 06 - Examining Alternative Explanations for Red Covid'
subtitle: "How to have an argument with regression"
author: "Your Group Members Names Here"
date: "Last Updated `r format(Sys.Date())`"
format:
  html:
    toc: true
    toc-location: right
    toc-float: true
    toc-depth: 2
    number-sections: true
    embed-resources: true
execute: 
  eval: true
  echo: true
  warning: false
  message: false
---


# Overview {.unnumbered}

Today we will explore the critiques and alternative explanations for the phenomena of "Red Covid" discussed by the *NYT*'s David Leonhardt in articles last fall [here](https://www.nytimes.com/2021/09/27/briefing/covid-red-states-vaccinations.html) and more recently [here](https://www.nytimes.com/2021/09/27/briefing/covid-red-states-vaccinations.html).

Recall the core thesis of Red Covid is something like the following:

> Since Covid-19 vaccines became widely available to the general public in the spring of 2021, Republicans have been less likely to get the vaccine. Lower rates of vaccination among Republicans have in turn led to higher rates of death from Covid-19 in Red States compared to Blue States.

A skeptic of this claim might argue that relationship between electoral and epidemelogical outcomes is *spurious*, saying somthing like:

> There are lots of ways that Red States differ from Blue States --- demographics, economics, geography, culture, and so on -- and it is these differences that explain the phenomena of Red Covid. If we were to control for these *omitted variables* the relationship between a state's partisan leanings and Covid-19 would go away.

In this lab, we will see how we can explore these claims using **multiple regression** to **control** for competing explanations.

To accomplish this we will:

1.  Get set up to work (10 minutes)

Then we will estimate and interpret a series of regression models:

2.  A baseline Red Covid model using simple bivariate regression using the `Republican vote share` of states to predict the 14-day average of per capita `Covid-19 deaths` on September 23, 2021 (10 Minutes)

3.  A multiple regression model controlling for `Republican vote share` the `median age` (15 minutes)

4.  A model controlling for `Republican vote share`, the `median age` **and** `median income` (15 minutes)

5.  A model controlling for `Republican vote share`, the `median age` `median income` **and** `vaccination rates` (15 minutes)

6.  A model using `Republican vote share`, the `median age` `median income` to predict `vaccination rates` (15 minutes)

Finally, we'll take the weekly survey which will serve as a mid semester check in.

One of these 6 tasks (excluding the weekly survey) will be randomly selected as the graded question for the lab.

You will work in your assigned groups. Only one member of each group needs to submit the html file of lab.

This lab **must** contain the names of the group members in attendance.

If you are attending remotely, you will submit your labs individually.

Here are your assigned groups for the semester.

```{r}
#| label: groups
#| echo: false
groups_df <- readr::read_csv("https://pols1600.paultesta.org/files/groups.csv")

DT::datatable(groups_df)
```

# Goals {.unnumbered}

The primary goal of this lab is to give you lots of practice estimating and interpreting multiple regression models.

Questions 2-6 will all ask you to do some combination of the following:

1.  **Fit** a linear model or models using `lm()`
2.  **Summarize** the results of this model using `summary()`
3.  **Present** the results of this model or models in a regression table
4.  **Interpret** your models substantively for other human beings

Question 2 will also give you practice visualizing bivariate relationships

In general, when we interpret the regression coefficients from a linear model we want to know their:

-   **Sign** is the coefficient positive or negative? As the predictor increases does the outcome tend to increase (positive) or decrease (negative)
-   **Size** is the change in the predicted value of the outcome **substantively** meaningful?
    -   To facilitate comparisons between predictors, we will use [standardized measures](https://pols1600.paultesta.org/slides/06-slides#/standardized-variables){target="_blank"} of Republican Vote Share, Median Age, and Median Income, and Percent Vaccinated
-   **Significance** today we'll put the cart before the horse, and say, coefficient that has one or more asterisks `*` next to it is *statistically significant*
    -   Over the next several weeks, we'll talk in great detail the logic and theory behind statistical inference
    -   For today, we'll do a bit of "star-gazing" using these asterisks[^1] as heuristics to help us evaluate the claims of Leonhardt and his skeptics.

[^1]: In short, these `*` correspond to $p-values$ below different thresholds. One `*` typically means $p < 0.05$. A p-value is a conditional probability that arises from a hypothesis test summarizing the likelihood of observing a particular test statistic (here a regression coefficient, or more specifically, a t-statistic which is the regression coefficient divided by its standard error) given a paritcular hypothesis (typically, but not allows a *null hypothesis* that the true coefficient is 0). In sum, a p-value assess the likelihood of seeing what we did, if in fact, there was no relationship. If that likelihood is small (p\<0.05), we reject the claim of no relationship. We remain uncertain about the true value of the coefficient, but we are pretty confident it's not 0.

# Workflow {.unnumbered}

# Please knit this .qmd file {.unnumbered}

As with every lab, you should:

-   Download the file
-   Save it in your course folder
-   **Update the `author:` section of the YAML header to include the names of your group members in attendance.**
-   Render the document
-   Open the html file in your browser (Easier to read)
-   Write your code in the **chunks provided under each section**
-   Comment out or delete any test code you do not need
-   **Render the document again after completing a section or chunk** (Error checking)
-   Upload the final lab to [Canvas](https://canvas.brown.edu/courses/1098409/assignments){target="_blank"}.

# Get setup to work

## Load Packages

First lets load the libraries we'll need for today.

There's one new package, `htmltools` which we'll use to display regression tables while we work.

```{r}
#| label: packages

the_packages <- c(
  ## R Markdown
  "kableExtra","DT","texreg","htmltools",
  ## Tidyverse
  "tidyverse", "lubridate", "forcats", "haven", "labelled",
  ## Extensions for ggplot
  "ggmap","ggrepel", "ggridges", "ggthemes", "ggpubr", 
  "GGally", "scales", "dagitty", "ggdag", "ggforce",
  # Data 
  "COVID19","maps","mapdata","qss","tidycensus", "dataverse",
  # Analysis
  "DeclareDesign", "easystats", "zoo"
)

# Define function to load packages
ipak <- function(pkg){
    new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
    if (length(new.pkg)) 
        install.packages(new.pkg, dependencies = TRUE)
    sapply(pkg, require, character.only = TRUE)
}

ipak(the_packages)
```

## Load the data

Next we'll load the data that we created in class on Tuesday which provides a snapshot of the state of Covid-19 on September 23, 2021 in the U.S.

```{r}
#| label: data
load(url("https://pols1600.paultesta.org/files/data/06_lab.rda"))
```

After running this code, the data frame `covid_lab` should appear in your environment pane in R Studio

## Describe the data

In the code chunk below, please write some code get an high level overview of the data:

```{r}
#| label: hlo
# High level overview
# Number of observations and variables


# Names of variables


# Glimpse of data


# Summary of data


# Calculate standard deviations



```

Please use this HLO to answer the following questions:

-   **How many observations are there:**

-   **What is an observation (i.e. what is the unit of analysis):**

-   **What is the primary outcome variable for today:**

-   **What are the four main predictors we'll be using:**

-   **Will we be using the the raw values of these predictors or their standardized values?**

-   **What is the standard deviation of our outcome and predictor variables:**

    -   Covid-19 deaths:
    -   Republican vote share:
    -   Median age:
    -   Median income:
    -   Vaccination Rate:

# Estimate and interpet the baseline Red Covid model

First let's estimate and interpret the following model:

$$\text{New Covid Deaths} = \beta_0 + \beta_1 \text{Republican Vote Share}_{std} + \epsilon$$

**All you have to do is run the code chunks below and then interpret the results**

When you visualize this model, you will have to **write comments in the code**

Then, you'll use this code as guide for subsequent sections.

## Fit the model

```{r}
#| label: m1
m1 <- lm(new_deaths_pc_14day ~ rep_voteshare_std, covid_lab)

```

## Summarize the results

Now we apply the `summary()` function to our model `m1`

```{r}
#| label: m1sum
summary(m1)
```

We see that m1 returns two coefficients, which define a line of best fit predicting Covid-19 deaths with the Republican vote share of the 2020 Presidential election:

-   $\beta_0$ corresponds to the intercept. This is model's prediction for a state where Trump got 0 percent of the vote. This is typically not something we care about.

-   $\beta_1$ corresponds to the slope. Because we used a standardized measure of vote share, we would say that a 1-standard deviation (about 10 percentage points) increase in Republican vote share is associated with a 0.23 increase the average number of new Covid-19 deaths. Given that this per-capita measure has a standard deviation of 0.4, this is a fairly sizable association.

-   Finally, note that last column of `summary(m1)` `Pr(>|t|)` both the coefficients for the intercept $(\beta_0)$ and `rep_voteshare_std` ($(\beta_1)$) are **statistically significant** (ie have an `*` next to them).

## Display the model as a regression table

Next we'll format the results of `summary(m1)` into a **regression table** using the `htmlreg()` function.

Regression tables are a the standard way of concisely presenting the results of regression models.

-   Each named row corresponds to the coefficients form the model

-   If there is an asterisks next to a coefficient, that coefficient is **statistically significant** with a p value below a certain threshold.

-   The numbers in parentheses below each coefficient correspond to the **standard error** of the coefficient (more on that later)[^2]

-   The bottom of the table contains summary statistics of of our model, which we'll ignore for today.

[^2]: A standard error is another one of those things in the cart that we're putting before the horse today. Briefly, it is an estimate of the standard deviation of the sampling distribution of a coefficient and describes how much our coefficient might vary had we had a different sample...

The code after `htmlreg(m1)` allows you to see what output of the table will look like in the html document while you're working in the qmd file.

```{r}
#| label: m1tab
#| echo: false
#| results: asis
htmlreg(list(m1))%>% HTML() %>% browsable()
```

## Visualize the model

Now let's visualize the results of our `m1` with a scatter plot.

In the code chunk below, I've written some comments to help you get started. You can also refer to last [week's lab](https://pols1600.paultesta.org/labs/comments/05-lab-comments#visualize-the-relationships-between-republican-vote-share-and-deaths-from-covid-19-on-september-23-2021){target="_blank"} for help



```{r}
#| label: m1fig

# 1. Tell ggplot what data to use

# 2. Set the aesthetic mappings of our figure
 
# 3. Draw points with x values corresponding to Rep vote share and y values corresponding to Covid deaths. 

# 4. Add labels using `label=state_po` aesthetic (set in aes()) above

# 5. Plot the regression model using geom_smooth(method = "lm")

# 6. Change the axis labels


```

## Interpret the results

**In a sentence our two, summarize the results of your analysis in this section**

You words here!

# Estimate a model controlling for age

Suppose a skeptic reading the *New York Times* took issue with Leonhardt's claims, and said what's really behind the claim of Red Covid is that Republican states tend to be older and older people are more at Risk of Dying from Covid-19.

One way we could address this critique is by estimating a multiple regression model that **controlled** for age.

$$\text{New Covid Deaths} = \beta_0 + \beta_1 \text{Repbulican Vote Share}_{std} + \beta_2 \text{Median Age}_{std} + \epsilon$$

If our skeptic is right, what should happen to coefficients:

-   **Republican Vote Share:**
    -   **Sign:** (Negative / Positive / Unclear)
    -   **Size:** (Increase compared `m1`/ Decrease compared `m1` / Unclear)
    -   **Significance:** (Significant / Non-Signficant)
-   **Median Age:**
    -   **Sign:** (Negative/Positive/Unclear)
    -   **Size:** (Large/Small/Unclear)
    -   **Significance:** (Significant / Non-Significant)

If our critique is right, and the relationship between Partisanship and Covid is confounded by the fact that Red States tend to be older and older people are more likely to die from Covid, then controlling for Age, the coefficient on Republican Vote share should decrease in size (get closer to 0) and lose significance, while the coefficient on Age should be positive (older states have more Covid-19 deaths) and statistically signficant. I'm not sure I have a good sense about the size or magnitude of this effect.

## Fit the model

Now let's test our skeptics' claims by fitting a model `m2` that controls for Age (`med_age_std`).

-   Remember the first argument in `lm()` is formula of the form `outcome variable ~ predictor1 + predictor2 + ...`

```{r}
#| label: m2
# m2
```

## Summarize the model

Now let's print out a statistical summary of `m2`

```{r}
#| label: m2sum
# summary of m2

```

## Display the model as a regression table

Next, let's create a regression table that displays `m1` in the first column and `m2` in the second column.

-   To do this, change `list(m1)` from the code above to `list(m1, m2)`

```{r}
#| label: m2tab
#| echo: false
# regression table of m1, m2

```

## Interpret the results

**In a few sentences, explain whether the results from `m2` support the skeptics criticisms or not?**

# Estimate a model controlling for age and income

Undeterred, our skeptic now argues that it's not just age that matters but also socioeconomic factors like wealth.

Let's test this claim using the following model:

$$\text{New Covid Deaths} = \beta_0 + \beta_1 \text{Repbulican Vote Share}_{std} + \beta_2 \text{Median Age}_{std} + \beta_3\text{Median Income}_{std}\epsilon$$ If the skeptic is right, then controlling for age and income, the coefficient on Republican vote share should **FILL IN WHAT SHOULD HAPPEN TO THIS COEFFICIENT**

## Fit the Model

Please fit a model called `m3` implied by the skeptic's revised claims

```{r}
#| label: m3
# m3

```

## Summarize the model

Summarize the model `m3` using `summary()`

```{r}
#| label: m3sum
# summary m3

```

## Display the models in a regression table

And then display the results of models `m1`, `m2`, and `m3`.

```{r}
#| label: m3tab
# regression table of m1, m2, m3

```

## Interpret the skeptic's claims

**In a few sentences, explain whether the results from `m3` support the skeptics criticisms or not?**



# Estimate a model controlling for age, income, and percent vaccinated

Hmm, maybe our skeptic has a point. Let's estimate a model that controls for everything from `m3` as well as the vaccination rate in each state.

$$\text{New Covid Deaths} = \beta_0 + \beta_1 \text{Rep Vote Share}_{std} + \beta_2 \text{Median Age}_{std} + \beta_3\text{Median Income}_{std}+\beta_4\text{Percent Vaxxed}_{std}\epsilon$$

## Fit the model

You know the drill.

```{r}
#| label: m4
# m4

```

## Summarize the results

Again, let's get a quick summary of our results

```{r}
#| label: m4sum
# summary of m4

```

## Display the models in a regression table

And add `m4` to list of models in our regression table

```{r}
#| label: m4tab
#| echo: false
# regression table of m1, m2, m3, m4

```

## Interpet the results

**Briefly interpret the results of `m4`**

# Model percent vaccinated in a state controlling for age, income, and republican vote share

Hmm, how should we make sense of `m4`. Let's fit one last model, that predicts vaccination rates as a function of Republican vote share, median age, and median income in a state.

$$\text{Percent Vaxxed} = \beta_0 + \beta_1 \text{Repbulican Vote Share}_{std} + \beta_2 \text{Median Age}_{std} + \beta_3\text{Median Income}_{std}\epsilon$$ Briefly, please describe whether we should expect a statistically significant positive or negative relationship with vaccination rates or whether we should expect no statistically significant relationship.

-   $\beta_1$ Republican Vote Share: (Negative / Positive / No Relationship)

-   $\beta_2$ Median Age: (Negative / Positive / No Relationship)

-   $\beta_3$ Median Income: (Negative / Positive / No Relationship)

## Fit the model

Now let's fit the model. For ease of interpretation, let's use the unstandardized measure of vaccination rates, `percent_vaccinated` as our outcome variable.

```{r}
#| label: m5
# m5

```

## Summarize the results

And summarize the results

```{r}
#| label: m5sum
# summary of m5

```

## Display the results in a regression table

Display them in a regression table

```{r}
#| label: m5tab
#| echo: false
# regression table for m5

```

## Interpret the results

**Summarize the results of `m5` and offer some broader discussion of what we've learned today**

# Take the Class Survey {.unnumbered}

Please take a few moments to complete the [class survey](https://brown.co1.qualtrics.com/jfe/form/SV_aVqL4lh92U7igGG){target="_blank"} for this week.
